---
title: "Penguins size comparisons"
author: "Jessica Madden"
date: "November 18, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms

ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

#QQ-plot compare our sample to if our data was perfectly normally distributed, asseses normality

ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

#Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g), 
    median_mass = median(body_mass_g), 
    sd_mass = sd(body_mass_g), 
    sample_size = n()
  )

```

Get vectors of body_mass_g for female and male chinstrap penguins seperately

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g) #gets just a vector of body mass values

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```


### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```

Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins?
- interested in one tail of the distribution

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater") #order of vectors matters now, insert alternative between vectors to ask question, one-sided t test
```

